---
title: "Football statistics dashboard"
output:
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
runtime: shiny
---

<style>
  .center-text {
    text-align: center;
  }
</style>

<div class="center-text">
  The best interactive dashboard on the market with statistics about football in the top 5 leagues
</div>


```{r setup, include = FALSE}
library(flexdashboard) 
library(tidyverse)
library(ggrepel)
library(DT)
library(shiny)
library(ggplot2)
#knitr::opts_chunk$set(fig.width = 5)


#read the data
data1 <- read_csv(file = "data/understat_per_game.csv")
data2 <- read_csv(file = "data/understat.com.csv")
#data3 <- read.csv(file = "data/matches.csv")

colnames_1 = c("league", "year", "stadium", "exp_goal", "exp_goal_against", 
               "no_pen_exp_goal", "no_pen_exp_goal_against", "deep", 
               "deep_allowed", "scored", "conceded", "exp_pts", "result", "date", 
               "wins", "draws", "loses", "pts", "no_pen_exp_goal_diff", "ppda_coef", 
               "ppda_att", "ppda_def", "oppda_coef", "oppda_att", "oppda_def", 
               "team", "exp_goal_diff", "exp_goal_against_diff", 
               "exp_pts_diff")

colnames(data1) = colnames_1

colnames_2 = c("league", "year", "position", "team", "matches", "wins", "draws", 
               "loses", "scored", "conceded", "pts", "exp_goal", "exp_goal_diff", 
               "no_pen_exp_goal", "exp_goal_against", "exp_goal_against_diff", 
               "no_pen_exp_goal_against", 
               "no_pen_exp_goal_diff", "ppda_coef", "oppda_coef", "deep", 
               "deep_allowed", "exp_pts", "exp_pts_diff")

colnames(data2) = colnames_2

data_per_game <- data1 |> 
  mutate("stadium" = as.factor(stadium), 
         "result" = as.factor(result), 
         "league" = as.factor(league),
         "pts" = as.factor(pts)) |> 
  filter(league != "RFPL")

data_per_club <- data2 |> 
  filter(league != "RFPL") |> 
  mutate("goal_diff" = scored - conceded)

## Check to see no duplicates and the pairing of opponent is done correctly on a 1-1 basis
length(data_per_game$league[duplicated(cbind(data_per_game$league, 
                                             data_per_game$year, 
                                             data_per_game$date, 
                                             data_per_game$scored, 
                                             data_per_game$conceded, 
                                             data_per_game$exp_goal_against, 
                                             data_per_game$exp_goal, 
                                             data_per_game$ppda_coef, 
                                             data_per_game$oppda_coef))])

## Link opponents to each respective match
temp_opponents = data_per_game |> select(league, year, date, scored, 
                                         conceded, exp_goal_against, exp_goal, 
                                         ppda_coef, oppda_coef, "opponent" = team)

data_per_game = data_per_game |> 
  left_join(temp_opponents, by = c("league" = "league", "year" = "year", 
                                   "date" = "date", "scored" = "conceded", 
                                   "conceded" = "scored", 
                                   "exp_goal_against" = "exp_goal", 
                                   "exp_goal" = "exp_goal_against", 
                                   "ppda_coef" = "oppda_coef", 
                                   "oppda_coef" = "ppda_coef"))


## Df created for plot 5 and 6 (Joao)
wins_home <- data_per_game |>
  filter(result =="w", stadium == "h") |>
  group_by(year, league, team) |>
  summarize(home_wins = n())

wins_away <- data_per_game |>
  filter(result =="w", stadium != "h") |>
  group_by(year, league, team) |>
  summarize(away_wins = n())

winjoin <- full_join(wins_home, wins_away, by = c("year", "league", "team"))

data_selected <- data_per_club |>
  select(year, league, team, position)

wins <- winjoin |>
  left_join(data_selected, by = c("year", "league", "team")) |>
  filter(position < 5)

# ## Df cleaning and created for plot 7 (Joao)
# games_data_filtred <- data3|> 
#   select(-home_starting_1_num:-away_starting_14)|>
#   select(-home_starting_15_num:-home_bench_15_minute)|>
#   select(-attendance:-league)|>
#   mutate("home_yellowCards"=as.numeric(home_yellowCards),
#          "away_yellowCards"= as.numeric(away_yellowCards),
#          "League" = as.factor(League))|>
#   filter(League != "Dutch Eredivisie")|>
#   filter(year < 2020 & year > 2013) |>
#   mutate(home = recode(home,
#                        "AFC Bournemouth" = "Bournemouth",
#                        "Brighton & Hove Albion" = "Brighton",
#                        "Cardiff City" = "Cardiff",
#                        "Huddersfield Town" = "Huddersfield",
#                        "Hull City" = "Hull",
#                        "Leicester City" = "Leicester",
#                        "Norwich City" = "Norwich",
#                        "Stoke City" = "Stoke",
#                        "Swansea City" = "Swansea",
#                        "Tottenham Hotspur" = "Tottenham",
#                        "West Ham United" = "West Ham",
#                        "FC Augsburg" = "Augsburg",
#                        "Borussia Monchengladbach" = "Borussia M.Gladbach",
#                        "SV Darmstadt 98" = "Darmstadt",
#                        "Fortuna Düsseldorf" = "Fortuna Duesseldorf",
#                        "SC Freiburg" = "Freiburg",
#                        "Hamburg SV" = "Hamburger SV",
#                        "TSG Hoffenheim" = "Hoffenheim",
#                        "FC Ingolstadt 04" = "Ingolstadt",
#                        "Mainz" = "Mainz 05",
#                        "Nurnberg" = "Nuernberg",
#                        "SC Paderborn 07" = "Paderborn",
#                        "RB Leipzig" = "RasenBallsport Leipzig",
#                        "1. FC Union Berlin" = "Union Berlin",
#                        "VfL Wolfsburg" = "Wolfsburg",
#                        "Almería" = "Almeria",
#                        "Alavés" = "Alaves",
#                        "Athletic Bilbao"="Athletic Club",
#                        "Córdoba" = "Cordoba",
#                        "Deportivo La Coruña" = "Deportivo La Coruna",
#                        "Granada CF" = "Granada",
#                        "Huesca" = "SD Huesca",
#                        "Leganés"="Leganes",
#                        "Málaga" = "Malaga",
#                        "Sevilla FC" = "Sevilla",
#                        "Sporting Gijón" = "Sporting Gijon",
#                        "AS Roma" = "Roma",
#                        "Chievo Verona" = "Chievo",
#                        "Internazionale" = "Inter",
#                        "Parma" = "Parma Calcio 1913",
#                        "US Pescara" = "Pescara",
#                        "SPAL" = "SPAL 2013",
#                        "Hellas Verona" = "Verona")) |>
#   mutate(away = recode(away,
#                        "AFC Bournemouth" = "Bournemouth",
#                        "Brighton & Hove Albion" = "Brighton",
#                        "Cardiff City" = "Cardiff",
#                        "Huddersfield Town" = "Huddersfield",
#                        "Hull City" = "Hull",
#                        "Leicester City" = "Leicester",
#                        "Norwich City" = "Norwich",
#                        "Stoke City" = "Stoke",
#                        "Swansea City" = "Swansea",
#                        "Tottenham Hotspur" = "Tottenham",
#                        "West Ham United" = "West Ham",
#                        "FC Augsburg" = "Augsburg",
#                        "Borussia Monchengladbach" = "Borussia M.Gladbach",
#                        "SV Darmstadt 98" = "Darmstadt",
#                        "Fortuna Düsseldorf" = "Fortuna Duesseldorf",
#                        "SC Freiburg" = "Freiburg",
#                        "Hamburg SV" = "Hamburger SV",
#                        "TSG Hoffenheim" = "Hoffenheim",
#                        "FC Ingolstadt 04" = "Ingolstadt",
#                        "Mainz" = "Mainz 05",
#                        "Nurnberg" = "Nuernberg",
#                        "SC Paderborn 07" = "Paderborn",
#                        "RB Leipzig" = "RasenBallsport Leipzig",
#                        "1. FC Union Berlin" = "Union Berlin",
#                        "VfL Wolfsburg" = "Wolfsburg",
#                        "Almería" = "Almeria",
#                        "Alavés" = "Alaves",
#                        "Athletic Bilbao"="Athletic Club",
#                        "Córdoba" = "Cordoba",
#                        "Deportivo La Coruña" = "Deportivo La Coruna",
#                        "Granada CF" = "Granada",
#                        "Huesca" = "SD Huesca",
#                        "Leganés"="Leganes",
#                        "Málaga" = "Malaga",
#                        "Sevilla FC" = "Sevilla",
#                        "Sporting Gijón" = "Sporting Gijon",
#                        "AS Roma" = "Roma",
#                        "Chievo Verona" = "Chievo",
#                        "Internazionale" = "Inter",
#                        "Parma" = "Parma Calcio 1913",
#                        "US Pescara" = "Pescara",
#                        "SPAL" = "SPAL 2013",
#                        "Hellas Verona" = "Verona")) 
# 
# games_data_filtred <-rename(games_data_filtred, "league" = "League")
# 
# 
# #Remove the 98th row because it contains NA values (ManCity in 2014)
# row_to_remove <- which(games_data_filtred$id == "395672") 
# games_data <- games_data_filtred[-row_to_remove, ]
# 
# cards_home <- games_data |>
#   group_by(year, league, home) |>
#   summarize(Home_Total_yellowCards = sum(home_yellowCards)) |> 
#   rename("team" = home)
# 
# cards_away <- games_data |>
#   group_by(year,league, away) |>
#   summarize(Away_Total_yellowCards = sum(away_yellowCards)) |>
#   rename("team" = away)
# 
# #Regroup the 2 df EPL_cards_home & EPL_cards_away 
# cards <-cards_home |>
#   inner_join(cards_away, by = c("year","league", "team")) |>
#   mutate(league = recode(league,
#                          "English Premier League"="EPL",
#                          "German Bundesliga"="Bundesliga",
#                          "Spanish La Liga"="La_liga",
#                          "French Ligue 1"="Ligue_1",
#                          "Italian Serie A"="Serie_A"
#                           ))
# 
# #Incorporate the ranking of the team of each year
# cards_final <- cards |>
#   inner_join(data_per_club, by = c("year","league", "team")) |>
#   select(-wins:-goal_diff) |>
#   mutate(Total_yellowCards = Home_Total_yellowCards + Away_Total_yellowCards,
#          Total_yellowCards_per_game = Total_yellowCards/matches,
#          Home_Total_yellowCards_per_game = Home_Total_yellowCards/matches,
#          Away_Total_yellowCards_per_game = Away_Total_yellowCards/matches)
```





Team {data-icon="fa-futbol" data-orientation=rows}
==========================


Options {.sidebar}
------------------------

```{r}

selectInput("team_league", "Choose a competition:",
                c("Bundesliga" = "Bundesliga",
                  "Premier League" = "EPL",
                  "Serie A" = "Serie_A",
                  "La Liga" = "La_liga", 
                  "Ligue 1" = "Ligue_1" 
                  ))

selectInput("team_year", "Choose a season:",
                c("2014-2015" = "2014",
                  "2015-2016" = "2015",
                  "2016-2017" = "2016",
                  "2017-2018" = "2017", 
                  "2018-2019" = "2018", 
                  "2019-2020" = "2019"))

teams = reactive({sort(filter(data_per_club, league == input$team_league, year == input$team_year)$team)})

teams_opp = reactive({sort(filter(data_per_club, league == input$team_league, year == input$team_year, team != input$team_team)$team)})

renderUI({selectInput("team_team", "Choose a team:", teams())})

renderUI({selectInput("team_opp", 
                      "Choose a team for comparison:", 
                      teams_opp())})

```


tabset
-----------------------------------

<!-- ### tab1  -->

<!-- Here we can see the main statistics for the team (insert team name) and their historic results against team (insert opponents name) -->

<!-- row -->
<!-- --------------------------- -->

### Final Position
```{r fig.width=1, fig.height=0.5}

renderValueBox({
  position = filter(data_per_club, 
                    league == input$team_league, 
                    year == input$team_year, 
                    team == input$team_team)$position[1]
  
  box_icon = ""
  
  if(position == 1){
    box_icon = "ion-trophy"
  } else if(position <= 3){
    box_icon = "ion-ribbon-a"
  } else if(position <= 10){
    box_icon = "ion-android-happy"
  } else{
    box_icon = "ion-android-sad"
  }
  
  valueBox(value = position, icon = box_icon)
})

```

### Matches Won at Home
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "h")$wins)
  matches_won = sum(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "h")$wins)
  
  gauge(matches_won, min = 0, max = matches_played, gaugeSectors(
  success = c(2*matches_played/3, matches_played), warning = c(matches_played/3, 2*matches_played/3), danger = c(0, matches_played/3), colors = c("mediumseagreen", "palegoldenrod", "lightcoral")))
})

```
> Number of wins playing at home

### Matches Won Away
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "a")$wins)
  matches_won = sum(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "a")$wins)
  
  gauge(matches_won, min = 0, max = matches_played, gaugeSectors(
  success = c(2*matches_played/3, matches_played), warning = c(matches_played/3, 2*matches_played/3), danger = c(0, matches_played/3), colors = c("mediumseagreen", "palegoldenrod", "lightcoral")))
})

```
> Number of wins playing away

### Matches Lost at Home
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "h")$loses)
  matches_lost = sum(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "h")$loses)
  
  gauge(matches_lost, min = 0, max = matches_played, gaugeSectors(
  success = c(2*matches_played/3, matches_played), warning = c(matches_played/3, 2*matches_played/3), danger = c(0, matches_played/3), colors = c("lightcoral", "palegoldenrod", "mediumseagreen")))
})

```
> Number of loses playing at home

### Matches Lost Away
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "a")$loses)
  matches_lost = sum(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "a")$loses)
  
  gauge(matches_lost, min = 0, max = matches_played, gaugeSectors(
  success = c(2*matches_played/3, matches_played), warning = c(matches_played/3, 2*matches_played/3), danger = c(0, matches_played/3), colors = c("lightcoral", "palegoldenrod", "mediumseagreen")))
})

```
> Number of loses playing away

row
--------------------------

### Season Results

```{r}

renderDataTable({data_per_game |> filter(league == input$team_league, 
                                     team == input$team_team, 
                                     year == input$team_year) |> 
    mutate(date = format(date, "%Y-%m-%d")) |> 
    mutate("home_team" = ifelse(stadium == "h", team, opponent)) |> 
    mutate("away_team" = ifelse(stadium == "a", team, opponent)) |> 
    mutate("home_goals" = ifelse(stadium == "h", scored, conceded)) |> 
    mutate("away_goals" = ifelse(stadium == "a", scored, conceded)) |> 
    select(date, home_team, home_goals, away_goals, away_team) |> 
    rename("Date" = "date", 
           "Home Team" = "home_team", 
           "Home Goals" = "home_goals", 
           "Away Goals" = "away_goals", 
           "Away Team" = "away_team")})

```
> Full team's results for the season

row
----------------------------------

### Head to Head 

```{r}

renderTable({data_per_game |> filter(league == input$team_league, 
                                     team == input$team_team, 
                                     opponent == input$team_opp) |> 
    mutate(date = format(date, "%Y-%m-%d")) |> 
    mutate("home_team" = ifelse(stadium == "h", team, opponent)) |> 
    mutate("away_team" = ifelse(stadium == "a", team, opponent)) |> 
    mutate("home_goals" = ifelse(stadium == "h", scored, conceded)) |> 
    mutate("away_goals" = ifelse(stadium == "a", scored, conceded)) |> 
    select(date, home_team, home_goals, away_goals, away_team) |> 
    rename("Date" = "date", 
           "Home Team" = "home_team", 
           "Home Goals" = "home_goals", 
           "Away Goals" = "away_goals", 
           "Away Team" = "away_team")}, digits = 0, align = "c")
  

```
> Historic results between the two teams in the competition

### Cummulative points comparison per match day

```{r}

renderPlot({data_per_game |> 
    filter((team == input$team_team | team == input$team_opp) & 
             year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    group_by(team, match) |> 
    summarize(pts = sum(pts)) |> 
    mutate(pts_cum = cumsum(pts)) |> 
    ggplot(mapping = aes(match, pts_cum, color = team)) + 
    geom_line() + 
    geom_point() + 
    labs(x = "Match Day", y = "Cummulative Points") + 
    theme_bw()})

```
> Comparison of performance of both teams throughout the season

row {.tabset .tabset-fade}
--------------------------

### Expected Goals and Actual Goals Scored throughout the Season

```{r}

renderPlot({data_per_game |> 
    filter(league == input$team_league, 
           team == input$team_team, 
           year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    ggplot() + 
    geom_line(mapping = aes(match, exp_goal, color = "Expected Goals"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, exp_goal, color = "Expected Goals"), 
              show.legend = TRUE) + 
    geom_line(mapping = aes(match, scored, color = "Goals Scored"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, scored, color = "Goals Scored"), 
              show.legend = TRUE) + 
    scale_color_manual("", breaks = c("Expected Goals", "Goals Scored"), 
                       values = c("red", "blue")) + 
    labs(x = "Match Day", y = "Goals / Expected Goals") + 
    theme_bw()})

```

### Expected Goals Against and Actual Goals Conceded throughout the Season

```{r}

renderPlot({data_per_game |> 
    filter(league == input$team_league, 
           team == input$team_team, 
           year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    ggplot() + 
    geom_line(mapping = aes(match, exp_goal_against, 
                            color = "Expected Goals Against"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, exp_goal_against, 
                            color = "Expected Goals Against"), 
              show.legend = TRUE) + 
    geom_line(mapping = aes(match, conceded, color = "Goals Conceded"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, conceded, color = "Goals Conceded"), 
              show.legend = TRUE) + 
    scale_color_manual("", breaks = c("Expected Goals Against", 
                                                    "Goals Conceded"), 
                       values = c("red", "blue")) + 
    labs(x = "Match Day", y = "Goals / Expected Goals") + 
    theme_bw()})

```



National League {data-icon="fa-shield"}
==============================

Options {.sidebar}
------------------------

```{r}
selectInput("league_league", "Choose a league:",
                c("Bundesliga" = "Bundesliga",
                  "Premier League" = "EPL",
                  "Serie A" = "Serie_A",
                  "La Liga" = "La_liga", 
                  "Ligue 1" = "Ligue_1"))
```


```{r}
selectInput("league_year", "Choose the season:",
                c("2014/2015" = 2014,
                  "2015/2016" = 2015,
                  "2016/2017" = 2016,
                  "2017/2018" = 2017, 
                  "2018/2019" = 2018,
                  "2019/2020" = 2019))
```




column
----------------------------


### Difference of goals per ranking position
```{r}
renderPlot({
  data_per_club |> 
    filter(league == input$league_league, year == input$league_year) |>
  ggplot() +
  geom_line(aes(x=position, y=goal_diff, color = "red" )) +
  geom_point(aes(x=position, y=goal_diff, color = "red")) +
  geom_line(aes(x=position, y=no_pen_exp_goal_diff, color = "green")) +
  geom_point(aes(x=position, y=no_pen_exp_goal_diff, color = "green")) +
  scale_y_continuous(breaks = seq(-70,100, by = 5)) +
  scale_x_continuous(breaks = seq(1,20,  by = 1)) +
  theme_linedraw() +
  labs(title = "Difference of goals per ranking position",
       x = "Ranking",
       y = "Difference of goals",
       caption = "Green curve represents expected difference of goals") + 
  theme(plot.title = element_text(hjust = 0.5))
})
```





### Link between goals difference and ranking


```{r}
custom_colors <- c("2014"= "darkgreen","2015"="darkseagreen","2016"="khaki","2017"="aquamarine","2018"="hotpink","2019"="deepskyblue")

renderPlot({
  data_per_club |>
    filter(league == input$league_league) |>
  ggplot(aes(x=position, y = goal_diff, color = as.factor(year), group = as.factor(year))) +
  geom_smooth(se = FALSE) +
  scale_color_manual(values = custom_colors) +
  scale_x_continuous(breaks = seq(1,20, by = 1)) +
  scale_y_continuous(breaks = seq(-60,80, by = 10)) +
  theme_bw() +
  labs(title = "Relation between goals difference per ranking",
       x = "Ranking",
       y = "Goals difference",
       color = "Years") + 
  theme(plot.title = element_text(hjust = 0.5))
})

```


column  {.tabset}
-------------------------------

### Plot of average goals scored per game and total wins

```{r}

renderPlot({
  
  data_per_club |>
              filter(league == input$league_league, year == input$league_year) |>
              mutate(Average_goalscored_per_game = scored/matches,
                     Average_goalsconceded_per_game = conceded/matches) |>
  ggplot(aes(x=wins,y=Average_goalscored_per_game)) +
  geom_point(aes(color = team)) +
  geom_point(size = 2.5, shape = 1) +
  geom_smooth(se=FALSE) +
  scale_x_continuous(breaks = seq(0,38, by = 1)) +
  scale_y_continuous(breaks = seq(0.50,3, by = 0.25 )) +
  theme_bw() +
  labs(title = "Average goals scored per game",
       x = "Games won",
       y = "Average goals scored per game",
       color = "Teams",
       caption = "The game isn't necessarily won") +
  ggrepel::geom_label_repel(aes(label=position), max.overlaps = Inf) + 
  theme(plot.title = element_text(hjust = 0.5))
  
})



```


### Plot of average goals conceded and total losses

```{r}

renderPlot({
  
  data_per_club |>
              filter(league == input$league_league, year == input$league_year) |>
              mutate(Average_goalscored_per_game = scored/matches,
                     Average_goalsconceded_per_game = conceded/matches) |>
  ggplot(aes(x=loses, y=Average_goalsconceded_per_game)) +
  geom_point(aes(color = team)) +
  geom_point(size = 2.5, shape = 1) +
  geom_smooth(se=FALSE) +
  scale_x_reverse(breaks = seq(0,30,1)) +
  scale_y_continuous(breaks = seq(0.50,3, by = 0.25 )) +
  theme_bw() +
  labs(title = "Average goals conceded per game",
       x = "Games lost",
       y = "Average goals conceded per game",
       color = "Teams",
       caption = "The game isn't necessarily lost") +
  ggrepel::geom_label_repel(aes(label=position), max.overlaps = Inf) + 
  theme(plot.title = element_text(hjust = 0.5))

})
```

### Link between wins at home and ranking

```{r}
renderPlot({
  wins |>
        filter(league == input$league_league) |>
  ggplot(aes(x=position, y=home_wins, fill = team)) +
  geom_bar(stat = "identity", alpha = 0.75) +
  scale_y_continuous(breaks = seq(0,18, by=3)) +
  theme_bw() +
  labs(title = "Relation between nb of wins at home and the ranking",
       x = "Ranking in the league",
       y = "Total wins at home",
       fill = "The top 4") +
  coord_flip() +
  facet_wrap(~year)

})
```

### Link between wins away and ranking

```{r}
renderPlot({
  wins |>
        filter(league == input$league_league) |>
ggplot(aes(x=position, y=away_wins, fill=team)) +
  geom_bar(stat="identity", alpha = 0.75) +
  scale_y_continuous(breaks = seq(0,18, by=3)) +
  theme_bw() +
  labs(title = "Relation between number of wins away and the ranking",
       x = "Ranking in the league",
       y = "Total of wins away",
       fill = "The top 4") + 
    coord_flip() + 
    facet_wrap(~ year)

})
```

<!-- ### Average yellow cards received per ranking -->

```{r}
# renderPlot({
#   cards_final |>
#                filter(league == input$league_league, year == input$league_year) |>
# ggplot() +
#   geom_col(aes(x=position, y=Total_yellowCards_per_game), size = 1, color = "darkblue", fill = "white") +
#   geom_line(aes(x=position, y= Total_yellowCards_per_game), size = 1.5, color = "red", linetype = "solid") +
#   geom_line(aes(x=position, y= Home_Total_yellowCards_per_game), size = 1.5, color = "darkgreen", linetype = "solid") +
#   geom_line(aes(x=position, y= Away_Total_yellowCards_per_game), size = 1.5, color = "dodgerblue", linetype = "solid") +
#   scale_y_continuous(breaks = seq(0,3, by = 0.10)) +
#   scale_x_continuous(breaks = seq(1,20,  by = 1)) +
#   theme_bw() +
#   labs(title = "Distribution of yellow cards received in total per game",
#        x = "Ranking",
#        y = "Total of yellow cards per game",
#        caption = "Blue curve represents yellow cards received per game away",
#        "Green curve represents yellow cards received per game at home") + 
#   theme(plot.title = element_text(hjust = 0.5))
# 
# })


```



Comparison across leagues {data-orientation=rows data-icon="fa-globe"}
=================================


Options {.sidebar}
------------------------

```{r}
# creating the variable that stores the value choosen by the user
selectInput("year_season", "Choose the season:",
                c("2014/2015" = 2014,
                  "2015/2016" = 2015,
                  "2016/2017" = 2016,
                  "2017/2018" = 2017, 
                  "2018/2019" = 2018,
                  "2019/2020" = 2019))
```



```{r}

# Define the checkBox with the different leagues that the user wants to be shown on the page

checkboxGroupInput("select_options", "Select the leagues:", 
                   choices = c("Bundesliga", "EPL", "Serie_A","La_liga" ,"Ligue_1"),
                   selected = c("Bundesliga", "EPL", "Serie_A","La_liga" ,"Ligue_1"))


```


row  {data-orentation=columns data-height=250}
---------------------------------

### Winning teams for each league given the choosen year 

```{r, fig.width=3}

# Printing on the dashboard the output of the table with the five winning teams in the leagues

renderTable({
      data_per_club |> filter(year == input$year_season, position == 1) |> 
      select(year,league, team:pts, exp_pts, exp_goal) }, digits = 0, align = "l", spacing = "s")

```


row 
----------------------------------

### Distribution of gol scored and total expected goal for each league

```{r}

# boxplot that compares the distribution of total goals scored with the total expected goals scored for all the years
 renderPlot({
      data_per_club |> filter(league %in% input$select_options) |> 
                                 ggplot() + 
                                 geom_boxplot(aes(league, exp_goal, color = "Exp Tot Goals" )) +
                                 geom_boxplot(aes(league, scored, color = "Tot goals scored")) +
                                 labs(x = "League",
                                      y = "Distribution of total goals scored") +
                                 scale_colour_manual("Exp goal per match", 
                                                     breaks = c("Exp Tot Goals", 
                                                                "Tot goals scored"), 
                                                     values = c("black","lightblue")) +
                                 facet_wrap(~ year) +
                                 theme_bw()
   
   
      })


```

> Looking at the distribution of total goals scored and total expected goals scored by the teams in the league help us to understand how equilibrate are their level of quality.

row
-------------------------------

### Difference goal scored for the first 10th ranked teams within each league

```{r, fig.height=3}

# Formatting the column of the match's date with just the year, month and day
data_per_game <-data_per_game |> mutate("date" = as.Date(date, "%Y-%m-%d")) 

# add a new column that shows for each match which was the match-day during the season
data_per_game1 <- data_per_game |> 
                  group_by(year, league, team) |> 
                  mutate("match_day" = row_number())


renderPlot({
  
  
# comparing the difference of goal scored for the first 10th ranked teams in each of the league chosen by user 
  
  data_per_club |> filter(year == input$year_season, 
                          position <= 10, 
                          league %in% input$select_options) |> 
                     ggplot() + 
                     geom_line(aes(position, goal_diff, color = league)) + 
                     labs(x = "Teams ranked by final position",
                          y = "Difference between goals scored and conceded ",
                          color = "Top 5 leagues") +
                     geom_point(aes(position, goal_diff, color = league)) +
                     scale_x_continuous(breaks = seq(1,10, by = 2)) +
                     theme_bw()
      
  })                                  


```


### Average teams' pressure within each league


```{r}


# BarPlot displays the average pressure in the opponent pitch by league, a measure of intensity during a game 

renderPlot({
  data_per_club |> filter(year == input$year_season, 
                          league %in% input$select_options) |> 
                 group_by(league) |> summarise(avg_opp_pressure = mean(ppda_coef)) |>
                 ggplot() + 
                 geom_bar(aes(league, avg_opp_pressure), stat = "Identity") + 
                 labs(x = "Top five football leagues",
                      y = "Average teams' pressure per league ") +
                 theme_bw()
  
})
```

> The average attack pressure in the opponent pitch gives us an estimator of the intensity of play 

row {.tabset .tabset-fade}
---------------------------------

### Average expected goal and goal scored per game played in the Premier League

```{r}

# for each match-day played in the Premier League the barplot and the line chart compare the average goal scored per match-day 
# with the average expected goal per match-day

renderPlot({
  
  data_per_game1 |> filter(year == input$year_season, 
                           league == "EPL") |> group_by(match_day) |> 
                  summarise(avg_goal_per_day = sum(scored)/(n()/2), 
                            avg_exp_goal = sum(exp_goal)/(n()/2)) |>
                    ggplot() + 
                  geom_bar(aes(match_day, avg_goal_per_day), fill="grey", stat = "Identity") +    
                  geom_line(aes(match_day, avg_exp_goal, color = "Exp Goals per match"), 
                                       linewidth = 0.5, 
                                       linetype = "dashed", show.legend = TRUE) + 
                  labs(x = "Match day",
                       y = "Average goal scored per match day") + 
                  scale_colour_manual("Exp goal per match", breaks = "Exp Goals per match", values = "red") +
                  scale_x_continuous(breaks = seq(1, 38, by = 2)) +
                  theme_bw()
  
  
})

```
> Here we can see for each match-day the difference between expected goal and gol scored for the Premier League, 


### Average goal scored per match played in the top 5 leagues

```{r}

# here is compared the average goal scored per match-day among leagues, the function geom_smooth emphasize the trend of this metric helping to visualize easily the pattern

renderPlot({
  
  data_per_game1 |> filter(year == input$year_season, 
                           league %in% input$select_options) |> group_by(league, match_day) |> 
                    summarise(avg_goal_per_day = sum(scored)/(n()/2)) |>
               ggplot() + 
               geom_line(aes(match_day, 
                             avg_goal_per_day, 
                             color = league), 
                             linetype = 2) +
               geom_smooth(aes(match_day, 
                               avg_goal_per_day, 
                               color = league), se = FALSE) +
               labs(x = "Match Day",
                    y = "Average goal scored per match day") +
               theme_bw()
})

```
> Approaching the end of the season the number of goals scored in a match on avearge increase exponentially


### Average expected goal per game played across the years 

```{r}

# here is the same plat as above but the expected goal is used and we show them for all of the years. here we want to compared if there has been a change in trend across different years

renderPlot({
  
  data_per_game1 |> filter(league %in% input$select_options) |> 
                    group_by(year, league, match_day) |>
                    summarise(avg_exp_goal = sum(exp_goal)/(n()/2)) |> 
                    ggplot() + geom_line(aes(match_day, 
                                             avg_exp_goal,
                                             color = league), 
                                         linetype = 6, 
                                         linewidth = 0.3) +
                    geom_smooth(aes(match_day, 
                                    avg_exp_goal, 
                                    color = league), 
                                se = FALSE) +
                    labs(x = "Match day",
                         y = "Average expected goal scored per match day") +
                    theme_bw() +
                    facet_wrap(~ year)
  })

              
```

> Along the years the expected goal per match-played increased torward the end of the season, supporting the idea that teams create more goal occasions.

###  Average passeges completed per goal scored during a match among the top 5 leagues

```{r}

# this Scatterplot compares the average of passages completed per goal with the average of goals per match played
# here again the trend line is shown to get an idea on the type of the trend.

renderPlot({
   
  data_per_club |> filter(year == input$year_season, 
                          league %in% input$select_options) |> group_by(league) |>
                   ggplot() + 
                   geom_point(aes(scored/matches, deep/matches)) +
                   geom_smooth(aes(scored/matches, deep/matches)) + 
                   labs(x = "Average goal scored per match played by the team",
                        y = "Average passages completed per match played by the team ") +
                   theme_bw() +
                   facet_wrap(~league) 
  
  })


```
> There is an increasing trend between passages completed before scoring a goal and number of goals, meaming that scoring more goals requires play more with the ball.


column
-------------------------

### Expected goals scored and goals scored per teams winning the competition 

```{r}
# for the five teams that won the league in their respective country is shown a barplot that compares their average tot goal scored with the expected total goals scored.

 renderPlot({
          
     data_per_club |> filter(position == 1, 
                             year == input$year_season, 
                             league %in% input$select_options) |> 
                      ggplot( aes(x = team)) + 
                      geom_bar(aes(y = exp_goal), position = "dodge", stat = "Identity") +
                      geom_point(aes(y = scored, color = "Goals scored"), shape = 17, size = 3) +
                      labs(x = "Winning teams for each league ",
                           y = "Expected total goal during the season ") +
                      scale_colour_manual("Total goals scored", breaks = "Goals scored", values = "red") +
                      theme_bw()
     
   })


```

> The winning teams scored more goals than the expected, indicating their ability to convert in goal shots


### Expected total points and effective points earned by the winning teams

```{r}

# the same as above but here we choose as metrics the total points earned and the total expected points

renderPlot({
          
     data_per_club |> filter(position == 1, 
                             year == input$year_season, 
                             league %in% input$select_options) |> 
                      ggplot( aes(x = team)) + 
                      geom_bar(aes(y = exp_pts), position = "dodge", stat = "Identity") +
                      geom_point(aes(y = pts, color = "Points"), shape = 17, size = 3) + 
                      labs(x = "Winning teams for the top 5 leagues",
                           y = "Expected total points earned during a season ") +
                      scale_colour_manual("Total points", breaks = "Points", values = "blue") +
                      theme_bw()
                     
   })
  
```

> Teams that won the league had a higher ability to bring at home the points compared to their expected points, meaning that their efficiency in defense and attack made the difference

