---
title: "Football statistics dashboard"
output:
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
runtime: shiny
---

<!-- centering the title of the dashboard that will be displayed on each page (Nicholas)-->

<style>
  .center-text {
    text-align: center;
  }
</style>

<div class="center-text">
  The best interactive dashboard on the market with statistics about football in the top 5 leagues
</div>


<!-- In this chunck there is all the code that is required to develop our dashboard and that is not going to be shown on the output. The whole design and structure (layout) of the dashboard was done by the same person. The content inside the dashboard was done by the whole team (one tab each member) (Nicholas) -->

```{r setup, include = FALSE}
library(flexdashboard) 
library(tidyverse)
library(ggrepel)
library(DT)
library(shiny)
library(ggplot2)
#knitr::opts_chunk$set(fig.width = 5)


#read the data (Nicholas)
data1 <- read_csv(file = "data/understat_per_game.csv")
data2 <- read_csv(file = "data/understat.com.csv")

colnames_1 = c("league", "year", "stadium", "exp_goal", "exp_goal_against", 
               "no_pen_exp_goal", "no_pen_exp_goal_against", "deep", 
               "deep_allowed", "scored", "conceded", "exp_pts", "result", "date", 
               "wins", "draws", "loses", "pts", "no_pen_exp_goal_diff", "ppda_coef", 
               "ppda_att", "ppda_def", "oppda_coef", "oppda_att", "oppda_def", 
               "team", "exp_goal_diff", "exp_goal_against_diff", 
               "exp_pts_diff")

colnames(data1) = colnames_1

colnames_2 = c("league", "year", "position", "team", "matches", "wins", "draws", 
               "loses", "scored", "conceded", "pts", "exp_goal", "exp_goal_diff", 
               "no_pen_exp_goal", "exp_goal_against", "exp_goal_against_diff", 
               "no_pen_exp_goal_against", 
               "no_pen_exp_goal_diff", "ppda_coef", "oppda_coef", "deep", 
               "deep_allowed", "exp_pts", "exp_pts_diff")

colnames(data2) = colnames_2

data_per_game <- data1 |> 
  mutate("stadium" = as.factor(stadium), 
         "result" = as.factor(result), 
         "league" = as.factor(league),
         "pts" = as.factor(pts)) |> 
  filter(league != "RFPL")

data_per_club <- data2 |> 
  filter(league != "RFPL") |> 
  mutate("goal_diff" = scored - conceded)



#Data Validation for data set per game (Santiago)


#Verify for NAs
sum(is.na(data_per_game))

#Check for possible typos on the league names
unique(data_per_game$league)

#Check for possible typos on the year the season began
unique(data_per_game$year)

#Check for possible typos on the home/away variable
unique(data_per_game$stadium)

#Summary of the expected goals + checking the extreme values
summary(data_per_game$exp_goal)
View(filter(data_per_game, exp_goal > 6))

#Summary of the no_pen_exp_goal variable
summary(data_per_game$no_pen_exp_goal)

#Summary of the deep variable
summary(data_per_game$deep)

#Summary of the goals scored + checking the extreme values
summary(data_per_game$scored)
View(filter(data_per_game, scored > 8))
#The values were verified to be correct

#Summary of the expected points
summary(data_per_game$exp_pts)

#Check for possible typos on the result variable
unique(data_per_game$result)

#Summary of the date
summary(data_per_game$date)

#Verifying there is no overcounting on wins
unique(data_per_game$wins)

#Verifying there is no overcounting on draws
unique(data_per_game$draws)

#Verifying there is no overcounting on loses
unique(data_per_game$loses)

#Verifying there is no apparent error on points
unique(data_per_game$pts)

#Summary of the no_pen_exp_goal_diff variable
summary(data_per_game$no_pen_exp_goal_diff)

#Summary of the ppda_coef variable
summary(data_per_game$ppda_coef)

#Verifying there is no typos in the names of the teams
sort(unique(data_per_game$team))

#Summary of the exp_goal_diff variable
summary(data_per_game$exp_goal_diff)

#Summary of the exp_pts_diff variable 
summary(data_per_game$exp_pts_diff)

#Verifying there is the correct number of teams for each league's seasons
data_per_game |> 
  select(league, year, team) |> 
  unique() |> 
  group_by(league, year) |> 
  summarize(number_of_teams = n()) |> 
  View()

#Verifying the information is complete (all the matches of all seasons)
data_per_game |> 
  group_by(league, year) |> 
  summarize(number_of_matches = n()/2) |> 
  View()
#We can see that there are 5 matches missing for the 2019/2020 Serie A.
#These matches were played on 02/08/20, and the dataset is updated until 01/08/20
#It's decided to keep the dashboard with this missing information, 
#awaiting for a future update to include the 5 extra matches.

#Verifying that no_pen_exp_goal is not greater than exp_goal
data_per_game |> 
  filter(no_pen_exp_goal > exp_goal) |> 
  View()

#Verifying that no_pen_exp_goal_against is not greater than exp_goal_against
data_per_game |> 
  filter(no_pen_exp_goal_against > exp_goal_against) |> 
  View()

#Verifying coherence between scores, results and points
data_per_game |> 
  filter(wins == 1, scored <= conceded) |> 
  View()

data_per_game |> 
  filter(draws == 1, scored != conceded) |> 
  View()

data_per_game |> 
  filter(loses == 1, scored >= conceded) |> 
  View()

data_per_game |> 
  filter(pts != wins*3 + draws*1) |> 
  View()

#Verifying the values of no_pen_exp_goal_diff
data_per_game |> 
  filter(no_pen_exp_goal_diff - (no_pen_exp_goal - no_pen_exp_goal_against) > 0.01) |> 
  View()

#Verifying the values of exp_goal_diff
data_per_game |> 
  filter(exp_goal_diff - (exp_goal - scored) > 0.01) |> 
  View()

#Verifying the values of exp_goal_against_diff
data_per_game |> 
  filter(exp_goal_against_diff - (exp_goal_against - conceded) > 0.01) |> 
  View()

#Verifying the values of exp_pts_diff
data_per_game |> 
  filter(exp_pts_diff - (exp_pts - as.numeric(as.character(pts))) > 0.01) |> 
  View()



#Data Validation for data set per club


#Verify for NAs
sum(is.na(data_per_club))

##Check for possible typos on the leagues' names
unique(data_per_club$league)

##Check for possible typos on the year
unique(data_per_club$year)

##Check for possible typos on the finishing position
unique(data_per_club$position)

##Check for possible typos on the teams' names
sort(unique(data_per_club$team))

##Check for possible errors on the number of matches played
unique(data_per_club$matches)
View(table(data_per_club$matches))
#The Bundesliga has 18 teams for a total of 34 matches per team.
#The other leagues have 20 teams, for a total of 38 matches per team.
#The 37 value accounts for the 10 teams of the 5 missing matches from 02/08/20
#The 27 and 28 values correspond to the 2019/20 Ligue 1 season, which did not play
#all the matches due to covid

#Summary of matches won
summary(data_per_club$wins)

#Summary of matches drawn
summary(data_per_club$draws)

#Summary of matches lost
summary(data_per_club$loses)

#Summary of scored goals
summary(data_per_club$scored)

#Summary of conceded goals
summary(data_per_club$conceded)

#Summary of pts
summary(data_per_club$pts)

#Summary of expected goals
summary(data_per_club$exp_goal)

#Summary of the exp_goal_diff variable
summary(data_per_club$exp_goal_diff)

#Summary of expected goals conceded
summary(data_per_club$exp_goal_against)

#Summary of the exp_goal_against_diff variable
summary(data_per_club$exp_goal_against_diff)

#Summary of the no_pen_exp_goal variable
summary(data_per_club$no_pen_exp_goal)

#Summary of the no_pen_exp_goal_against variable
summary(data_per_club$no_pen_exp_goal_against)

#Summary of the no_pen_exp_goal_diff variable
summary(data_per_club$no_pen_exp_goal_diff)

#Summary of the ppda_coef variable
summary(data_per_club$ppda_coef)

#Summary of the oppda_coef variable
summary(data_per_club$oppda_coef)

#Summary of the deep variable
summary(data_per_club$deep)

#Summary of the deep_allowed variable
summary(data_per_club$deep_allowed)

#Summary of the expected pts
summary(data_per_club$exp_pts)

#Summary of the exp_pts_diff variable
summary(data_per_club$exp_pts_diff)


#Verify there is no repeated position in a given season for any league
data_per_club |> 
  group_by(league, year, position) |> 
  summarise(frequency = n()) |> 
  filter(frequency != 1) |> 
  View()

#Verify the number of teams that has a determined number of matches in a season
data_per_club |> 
  group_by(league, year, matches) |> 
  summarise(frequency = n()) |> 
  View()
#The Bundesliga has 18 teams for a total of 34 matches per team.
#The other leagues have 20 teams, for a total of 38 matches per team.
#The 37 value accounts for the 10 teams of the 5 missing matches from 02/08/20
#The 27 and 28 values correspond to the 2019/20 Ligue 1 season, which did not play
#all the matches due to covid

#Verify that all the matches are counted in a result outcome
data_per_club |> 
  filter(matches - (wins + draws + loses) > 0.01) |> 
  View()

#Verify the points obtained are calculated properly
data_per_club |> 
  filter(pts - (wins*3 + draws*1) > 0.01) |> 
  View()

#Verifying the values of exp_goal_diff
data_per_club |> 
  filter(exp_goal_diff - (exp_goal - scored) > 0.01) |> 
  View()

#Verifying the values of exp_goal_against_diff
data_per_club |> 
  filter(exp_goal_against_diff - (exp_goal_against - conceded) > 0.01) |> 
  View()

#Verifying that no_pen_exp_goal is not greater than exp_goal
data_per_club |> 
  filter(no_pen_exp_goal > exp_goal) |> 
  View()

#Verifying that no_pen_exp_goal_against is not greater than exp_goal_against
data_per_club |> 
  filter(no_pen_exp_goal_against > exp_goal_against) |> 
  View()

#Verifying the values of no_pen_exp_goal_diff
data_per_club |> 
  filter(no_pen_exp_goal_diff - (no_pen_exp_goal - no_pen_exp_goal_against) > 0.01) |> 
  View()

#Verifying the values of exp_pts_diff
data_per_club |> 
  filter(exp_pts_diff - (exp_pts - pts) > 0.01) |> 
  View()

## Check to see no duplicates and the pairing of opponent is done correctly on a 1-1 basis (Santiago)
length(data_per_game$league[duplicated(cbind(data_per_game$league, 
                                             data_per_game$year, 
                                             data_per_game$date, 
                                             data_per_game$scored, 
                                             data_per_game$conceded, 
                                             data_per_game$exp_goal_against, 
                                             data_per_game$exp_goal, 
                                             data_per_game$ppda_coef, 
                                             data_per_game$oppda_coef))])

## Link opponents to each respective match (Santiago)
temp_opponents = data_per_game |> 
  select(league, year, date, scored, conceded, exp_goal_against, exp_goal, 
         ppda_coef, oppda_coef, "opponent" = team)

data_per_game = data_per_game |> 
  left_join(temp_opponents, by = c("league" = "league", "year" = "year", 
                                   "date" = "date", "scored" = "conceded", 
                                   "conceded" = "scored", 
                                   "exp_goal_against" = "exp_goal", 
                                   "exp_goal" = "exp_goal_against", 
                                   "ppda_coef" = "oppda_coef", 
                                   "oppda_coef" = "ppda_coef"))


## Df created for plot 5 and 6 (Joao)
wins_home <- data_per_game |>
  filter(result =="w", stadium == "h") |>
  group_by(year, league, team) |>
  summarize(home_wins = n())

wins_away <- data_per_game |>
  filter(result =="w", stadium != "h") |>
  group_by(year, league, team) |>
  summarize(away_wins = n())

winjoin <- full_join(wins_home, wins_away, by = c("year", "league", "team"))

data_selected <- data_per_club |>
  select(year, league, team, position)

wins <- winjoin |>
  left_join(data_selected, by = c("year", "league", "team")) |>
  filter(position < 5)

```



<!-- There are three main pages on our dashboard, here the first is defined. The line below define the orientation on how to add new plots and the icon that will represent this page. (Nicholas)-->

Team {data-icon="fa-futbol" data-orientation=rows}
==========================

<!-- this option allows to create a sidebar on the left of the page where the user can chooose and store his input. The option .sidebar create more pages with tabs (Nicholas)-->

Options {.sidebar}
------------------------

```{r}

#Defining the inputs for the Team tab (Santiago)
selectInput("team_league", "Choose a competition:",
                c("Bundesliga" = "Bundesliga",
                  "Premier League" = "EPL",
                  "Serie A" = "Serie_A",
                  "La Liga" = "La_liga", 
                  "Ligue 1" = "Ligue_1")
            )

selectInput("team_year", "Choose a season:",
                c("2014-2015" = "2014",
                  "2015-2016" = "2015",
                  "2016-2017" = "2016",
                  "2017-2018" = "2017", 
                  "2018-2019" = "2018", 
                  "2019-2020" = "2019")
            )

teams = reactive({sort(filter(data_per_club, 
                              league == input$team_league, 
                              year == input$team_year)$team)})

teams_opp = reactive({sort(filter(data_per_club, 
                                  league == input$team_league, 
                                  year == input$team_year, 
                                  team != input$team_team)$team)})

renderUI({
  selectInput("team_team", "Choose a team:", teams())
  })

renderUI({
  selectInput("team_opp", "Choose a team for comparison:", teams_opp())
  })

```


tabset
-----------------------------------


### Final Position
```{r fig.width=1, fig.height=0.5}

#Final position of the team for the season, with icon according to performance
#(Santiago)
renderValueBox({
  position = filter(data_per_club, 
                    league == input$team_league, 
                    year == input$team_year, 
                    team == input$team_team)$position[1]
  
  box_icon = ""
  
  if(position == 1){
    box_icon = "ion-trophy"
  } else if(position <= 3){
    box_icon = "ion-ribbon-a"
  } else if(position <= 10){
    box_icon = "ion-android-happy"
  } else{
    box_icon = "ion-android-sad"
  }
  
  valueBox(value = position, icon = box_icon)
})

```

### Matches Won at Home
```{r fig.width=1, fig.height=1}

#Rendering 4 Gauges that show the wins and looses of the team (home and away)
#as a proportion of the total matches played (Santiago)
renderGauge({
  matches_played = length(filter(data_per_game, 
                                 league == input$team_league, 
                                 year == input$team_year, 
                                 team == input$team_team, 
                                 stadium == "h")$wins)
  
  matches_won = sum(filter(data_per_game, 
                           league == input$team_league, 
                           year == input$team_year, 
                           team == input$team_team, 
                           stadium == "h")$wins)
  
  gauge(matches_won, min = 0, max = matches_played, 
        gaugeSectors(success = c(2*matches_played/3, matches_played), 
                     warning = c(matches_played/3, 2*matches_played/3), 
                     danger = c(0, matches_played/3), 
                     colors = c("mediumseagreen", "palegoldenrod", "lightcoral")))
})

```
> Number of wins playing at home

### Matches Won Away
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, 
                                 league == input$team_league, 
                                 year == input$team_year, 
                                 team == input$team_team, 
                                 stadium == "a")$wins)
  
  matches_won = sum(filter(data_per_game, 
                           league == input$team_league, 
                           year == input$team_year, 
                           team == input$team_team, 
                           stadium == "a")$wins)
  
  gauge(matches_won, min = 0, max = matches_played, 
        gaugeSectors(success = c(2*matches_played/3, matches_played), 
                     warning = c(matches_played/3, 2*matches_played/3), 
                     danger = c(0, matches_played/3), 
                     colors = c("mediumseagreen", "palegoldenrod", "lightcoral")))
})

```
> Number of wins playing away

### Matches Lost at Home
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, 
                                 league == input$team_league, 
                                 year == input$team_year, 
                                 team == input$team_team, 
                                 stadium == "h")$loses)
  
  matches_lost = sum(filter(data_per_game, 
                            league == input$team_league, 
                            year == input$team_year, 
                            team == input$team_team, 
                            stadium == "h")$loses)
  
  gauge(matches_lost, min = 0, max = matches_played, 
        gaugeSectors(success = c(2*matches_played/3, matches_played), 
                     warning = c(matches_played/3, 2*matches_played/3), 
                     danger = c(0, matches_played/3), 
                     colors = c("lightcoral", "palegoldenrod", "mediumseagreen")))
})

```
> Number of loses playing at home

### Matches Lost Away
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, 
                                 league == input$team_league, 
                                 year == input$team_year, 
                                 team == input$team_team, 
                                 stadium == "a")$loses)
  
  matches_lost = sum(filter(data_per_game, 
                            league == input$team_league, 
                            year == input$team_year, 
                            team == input$team_team, 
                            stadium == "a")$loses)
  
  gauge(matches_lost, min = 0, max = matches_played, 
        gaugeSectors(success = c(2*matches_played/3, matches_played), 
                     warning = c(matches_played/3, 2*matches_played/3), 
                     danger = c(0, matches_played/3), 
                     colors = c("lightcoral", "palegoldenrod", "mediumseagreen")))
})

```
> Number of loses playing away

row
--------------------------

### Season Results

```{r}

#Summary table with the results of the whole season for the team (Santiago)
renderDataTable({
  data_per_game |> filter(league == input$team_league, 
                          team == input$team_team, 
                          year == input$team_year) |> 
    mutate(date = format(date, "%Y-%m-%d")) |> 
    mutate("home_team" = ifelse(stadium == "h", team, opponent)) |> 
    mutate("away_team" = ifelse(stadium == "a", team, opponent)) |> 
    mutate("home_goals" = ifelse(stadium == "h", scored, conceded)) |> 
    mutate("away_goals" = ifelse(stadium == "a", scored, conceded)) |> 
    select(date, home_team, home_goals, away_goals, away_team) |> 
    rename("Date" = "date", "Home Team" = "home_team", 
           "Home Goals" = "home_goals", "Away Goals" = "away_goals", 
           "Away Team" = "away_team")
  })

```
> Full team's results for the season

row
----------------------------------

### Head to Head 

```{r}

#Historic results of matches between the two teams (Santiago)
renderTable({
  data_per_game |> filter(league == input$team_league, 
                          team == input$team_team, 
                          opponent == input$team_opp) |> 
    mutate(date = format(date, "%Y-%m-%d")) |> 
    mutate("home_team" = ifelse(stadium == "h", team, opponent)) |> 
    mutate("away_team" = ifelse(stadium == "a", team, opponent)) |> 
    mutate("home_goals" = ifelse(stadium == "h", scored, conceded)) |> 
    mutate("away_goals" = ifelse(stadium == "a", scored, conceded)) |> 
    select(date, home_team, home_goals, away_goals, away_team) |> 
    rename("Date" = "date", "Home Team" = "home_team", 
           "Home Goals" = "home_goals", "Away Goals" = "away_goals", 
           "Away Team" = "away_team")
  }, digits = 0, align = "c")

```
> Historic results between the two teams in the competition

### Cummulative points comparison per match day

```{r}

#Comparison of performance of both teams throughout the season (Santiago)
renderPlot({
  data_per_game |> 
    filter((team == input$team_team | team == input$team_opp) & 
             year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    group_by(team, match) |> 
    summarize(pts = sum(pts)) |> 
    mutate(pts_cum = cumsum(pts)) |> 
    ggplot(mapping = aes(match, pts_cum, color = team)) + 
    geom_line() + 
    geom_point() + 
    labs(x = "Match Day", y = "Cummulative Points") + 
    theme_bw()
  })

```
> Comparison of performance of both teams throughout the season

<!-- with the option .tabset-fade is possible to change the effect on changing a tab (Nicholas) -->

row {.tabset .tabset-fade}
--------------------------

### Expected Goals and Actual Goals Scored throughout the Season

```{r}

#Measure of offensive performance of the team throughout the season (Santiago)
renderPlot({
  data_per_game |> filter(league == input$team_league, 
                          team == input$team_team, 
                          year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    ggplot() + 
    geom_line(mapping = aes(match, exp_goal, color = "Expected Goals"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, exp_goal, color = "Expected Goals"), 
              show.legend = TRUE) + 
    geom_line(mapping = aes(match, scored, color = "Goals Scored"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, scored, color = "Goals Scored"), 
              show.legend = TRUE) + 
    scale_color_manual("", breaks = c("Expected Goals", "Goals Scored"), 
                       values = c("red", "blue")) + 
    labs(x = "Match Day", y = "Goals / Expected Goals") + 
    theme_bw()
  })

```

### Expected Goals Against and Actual Goals Conceded throughout the Season

```{r}

#Measure of defensive performance of the team throughout the season (Santiago)
renderPlot({data_per_game |> filter(league == input$team_league, 
                                    team == input$team_team, 
                                    year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    ggplot() + 
    geom_line(mapping = aes(match, exp_goal_against, 
                            color = "Expected Goals Against"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, exp_goal_against, 
                            color = "Expected Goals Against"), 
              show.legend = TRUE) + 
    geom_line(mapping = aes(match, conceded, color = "Goals Conceded"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, conceded, color = "Goals Conceded"), 
              show.legend = TRUE) + 
    scale_color_manual("", breaks = c("Expected Goals Against", 
                                                    "Goals Conceded"), 
                       values = c("red", "blue")) + 
    labs(x = "Match Day", y = "Goals / Expected Goals") + 
    theme_bw()
  })

```

<!-- the second page of the dashboard is create  (Nicholas)-->

National League {data-icon="fa-shield" data-orientation=rows}
==============================

Options {.sidebar}
------------------------

```{r}
selectInput("league_league", "Choose a league:",
                c("Bundesliga" = "Bundesliga",
                  "Premier League" = "EPL",
                  "Serie A" = "Serie_A",
                  "La Liga" = "La_liga", 
                  "Ligue 1" = "Ligue_1"))
```


```{r}
selectInput("league_year", "Choose the season:",
                c("2014/2015" = 2014,
                  "2015/2016" = 2015,
                  "2016/2017" = 2016,
                  "2017/2018" = 2017, 
                  "2018/2019" = 2018,
                  "2019/2020" = 2019))
```




row
----------------------------


### Difference of goals scored per team given the ranking position 
```{r}
renderPlot({
  data_per_club |> 
    filter(league == input$league_league, year == input$league_year) |>
    ggplot() +
    geom_line(aes(x=position, y=goal_diff, color = "Goals diff")) +
    geom_point(aes(x=position, y=goal_diff, color = "Goals diff")) +
    geom_line(aes(x=position, y=no_pen_exp_goal_diff, color = "Exp. goals diff")) +
    geom_point(aes(x=position, y=no_pen_exp_goal_diff, color = "Exp. goals diff")) +
    scale_y_continuous(breaks = seq(-70,100, by = 5)) +
    scale_x_continuous(breaks = seq(1,20,  by = 1)) +
    theme_bw() +
    labs(x = "Ranking position",
         y = "Difference of goals",
         color = "Graph legend",  #In order to change the colour name
         scale_colour_manual("Graph legend", 
                             breaks = c("Goals diff", 
                                        "Exp. goals diff"), 
                             values = c("red","green"))) +
    theme(plot.title = element_text(hjust = 0.5))
})
```

> As was predictable the best teams have the higher difference. It means that they performed better both in scoring and in the defense



### Link between goals difference and ranking


```{r}
custom_colors <- c("2014"= "darkgreen","2015"="darkseagreen","2016"="khaki","2017"="aquamarine","2018"="hotpink","2019"="deepskyblue")

renderPlot({
  data_per_club |>
    filter(league == input$league_league) |>
    ggplot(aes(x=position, y = goal_diff, color = as.factor(year), group = as.factor(year))) +
    geom_smooth(se = FALSE) +
    scale_color_manual(values = custom_colors) +
    scale_x_continuous(breaks = seq(1,20, by = 1)) +
    scale_y_continuous(breaks = seq(-60,80, by = 10)) +
    theme_bw() +
    labs(x = "Ranking position",
         y = "Difference of goals",
         color = "Years") + 
    theme(plot.title = element_text(hjust = 0.5))
  
})
```

> This plot tries to explain if among difference years there was a change in the difference goals scored per team along the ranking.  

row {.tabset .tabset-fade}
-------------------------------

### Goals scored per game and the total wins for each team


```{r}

renderPlot({
  
  data_per_club |>
    filter(league == input$league_league, year == input$league_year) |>
    mutate(Average_goalscored_per_game = scored/matches,
           Average_goalsconceded_per_game = conceded/matches) |>
    ggplot(aes(x=wins,y=Average_goalscored_per_game)) +
    geom_point(aes(color = team)) +
    geom_point(size = 2.5, shape = 1) +
    geom_smooth(se=FALSE) +
    scale_x_continuous(breaks = seq(0,38, by = 1)) +
    scale_y_continuous(breaks = seq(0.50,3, by = 0.25 )) +
    theme_bw() +
    labs(x = "Number of wins",
         y = "Goals scored per game",
         color = "Teams",
         caption = "The numbers inside the box are the ranking of the team") +
    ggrepel::geom_label_repel(aes(label=position), max.overlaps = Inf) + 
    theme(plot.title = element_text(hjust = 0.5))
  
})
```


### Goals conceded per game with the number of total losses for each team


```{r}

renderPlot({
  
  data_per_club |>
    filter(league == input$league_league, year == input$league_year) |>
    mutate(Average_goalscored_per_game = scored/matches,
           Average_goalsconceded_per_game = conceded/matches) |>
    ggplot(aes(x=loses, y=Average_goalsconceded_per_game)) +
    geom_point(aes(color = team)) +
    geom_point(size = 2.5, shape = 1) +
    geom_smooth(se=FALSE) +
    scale_x_reverse(breaks = seq(0,30,1)) +
    scale_y_continuous(breaks = seq(0.50,3, by = 0.25 )) +
    theme_bw() +
    labs(x = "Number of loses",
         y = "Goals conceded per game",
         color = "Teams",
         caption = "The numbers inside the box are the ranking of the team") +
    ggrepel::geom_label_repel(aes(label=position), max.overlaps = Inf) + 
    theme(plot.title = element_text(hjust = 0.5))
  
})
```

### Link between wins at home and ranking


```{r}
renderPlot({
  wins |>
    filter(league == input$league_league) |>
    ggplot(aes(x=position, y=home_wins, fill = team)) +
    geom_bar(stat = "identity", alpha = 0.75) +
    scale_y_continuous(breaks = seq(0,18, by=3)) +
    theme_bw() +
    labs(x = "Ranking in the league",
         y = "Number of wins at home",
         fill = "The top teams for each league") +
    theme(plot.title = element_text(hjust = 0.5)) +
    coord_flip() +
    facet_wrap(~year)

})
```

### Link between wins away and ranking

```{r}
  renderPlot({
    wins |>
      filter(league == input$league_league) |>
      ggplot(aes(x=position, y=away_wins, fill = team)) +
      geom_bar(stat = "identity", alpha = 0.75) +
      scale_y_continuous(breaks = seq(0,18, by=3)) +
      theme_bw() +
      labs(x = "Ranking in the league",
           y = "Number of wins away",
           fill = "The top 4 teams") +
      theme(plot.title = element_text(hjust = 0.5)) +
      coord_flip() +
      facet_wrap(~year)
   
})
```

<!-- creating the last page of the dashboard (Nicholas)-->

Comparison across leagues {data-orientation=rows data-icon="fa-globe"}
=================================


Options {.sidebar}
------------------------

```{r}
# creating the variable that stores the value choosen by the user (Nicholas)
selectInput("year_season", "Choose the season:",
                c("2014/2015" = 2014,
                  "2015/2016" = 2015,
                  "2016/2017" = 2016,
                  "2017/2018" = 2017, 
                  "2018/2019" = 2018,
                  "2019/2020" = 2019))
```



```{r}

# Define the checkBox with the different leagues that the user wants to be shown on the page (Nicholas)

checkboxGroupInput("select_options", "Select the leagues:", 
                   choices = c("Bundesliga", "EPL", "Serie_A","La_liga" ,"Ligue_1"),
                   selected = c("Bundesliga", "EPL", "Serie_A","La_liga" ,"Ligue_1"))


```


row  {data-orentation=columns data-height=250}
---------------------------------

### Winning teams for each league given the choosen year 

```{r, fig.width=3}

# Printing on the dashboard the output of the table with the five winning teams in the leagues (Nicholas)

renderTable({
      data_per_club |> filter(year == input$year_season, position == 1) |> 
      select(year,league, team:pts, exp_pts, exp_goal) }, digits = 0, align = "l", spacing = "s")

```


row 
----------------------------------

### Distribution of gol scored and total expected goal for each league

```{r}

# boxplot that compares the distribution of total goals scored with the total expected goals scored for all the years (Nicholas)
 renderPlot({
      data_per_club |> filter(league %in% input$select_options) |> 
                                 ggplot() + 
                                 geom_boxplot(aes(league, exp_goal, color = "Exp Tot Goals" )) +
                                 geom_boxplot(aes(league, scored, color = "Tot goals scored")) +
                                 labs(x = "League",
                                      y = "Distribution of total goals scored") +
                                 scale_colour_manual("Exp goal per match", 
                                                     breaks = c("Exp Tot Goals", 
                                                                "Tot goals scored"), 
                                                     values = c("black","lightblue")) +
                                 facet_wrap(~ year) +
                                 theme_bw()
   
   
      })


```

> Looking at the distribution of total goals scored and total expected goals scored by the teams in the league help us to understand how equilibrate are their level of quality.

row
-------------------------------

### Difference goal scored for the first 10th ranked teams within each league

```{r, fig.height=3}

# Formatting the column of the match's date with just the year, month and day (Nicholas)
data_per_game <-data_per_game |> mutate("date" = as.Date(date, "%Y-%m-%d")) 

# add a new column that shows for each match which was the match-day during the season (Nicholas)
data_per_game1 <- data_per_game |> 
                  group_by(year, league, team) |> 
                  mutate("match_day" = row_number())


renderPlot({
  
  
# comparing the difference of goal scored for the first 10th ranked teams in each of the league chosen by user (Nicholas)
  
  data_per_club |> filter(year == input$year_season, 
                          position <= 10, 
                          league %in% input$select_options) |> 
                     ggplot() + 
                     geom_line(aes(position, goal_diff, color = league)) + 
                     labs(x = "Teams ranked by final position",
                          y = "Difference between goals scored and conceded ",
                          color = "Top 5 leagues") +
                     geom_point(aes(position, goal_diff, color = league)) +
                     scale_x_continuous(breaks = seq(1,10, by = 2)) +
                     theme_bw()
      
  })                                  


```


### Average teams' pressure within each league


```{r}


# BarPlot displays the average pressure in the opponent pitch by league, a measure of intensity during a game (Nicholas)

renderPlot({
  data_per_club |> filter(year == input$year_season, 
                          league %in% input$select_options) |> 
                 group_by(league) |> summarise(avg_opp_pressure = mean(ppda_coef)) |>
                 ggplot() + 
                 geom_bar(aes(league, avg_opp_pressure), stat = "Identity") + 
                 labs(x = "Top five football leagues",
                      y = "Average teams' pressure per league ") +
                 theme_bw()
  
})
```

> The average attack pressure in the opponent pitch gives us an estimator of the intensity of play 

row {.tabset .tabset-fade}
---------------------------------

### Average expected goal and goal scored per game played in the Premier League

```{r}

# for each match-day played in the Premier League the barplot and the line chart compare the average goal scored per match-day 
# with the average expected goal per match-day (Nicholas)

renderPlot({
  
  data_per_game1 |> filter(year == input$year_season, 
                           league == "EPL") |> group_by(match_day) |> 
                  summarise(avg_goal_per_day = sum(scored)/(n()/2), 
                            avg_exp_goal = sum(exp_goal)/(n()/2)) |>
                    ggplot() + 
                  geom_bar(aes(match_day, avg_goal_per_day), fill="grey", stat = "Identity") +    
                  geom_line(aes(match_day, avg_exp_goal, color = "Exp Goals per match"), 
                                       linewidth = 0.5, 
                                       linetype = "dashed", show.legend = TRUE) + 
                  labs(x = "Match day",
                       y = "Average goal scored per match day") + 
                  scale_colour_manual("Exp goal per match", breaks = "Exp Goals per match", values = "red") +
                  scale_x_continuous(breaks = seq(1, 38, by = 2)) +
                  theme_bw()
  
  
})

```
> Here we can see for each match-day the difference between expected goal and gol scored for the Premier League, 


### Average goal scored per match played in the top 5 leagues

```{r}

# here is compared the average goal scored per match-day among leagues, the function geom_smooth emphasize the trend of this metric helping to visualize easily the pattern (Nicholas)

renderPlot({
  
  data_per_game1 |> filter(year == input$year_season, 
                           league %in% input$select_options) |> group_by(league, match_day) |> 
                    summarise(avg_goal_per_day = sum(scored)/(n()/2)) |>
               ggplot() + 
               geom_line(aes(match_day, 
                             avg_goal_per_day, 
                             color = league), 
                             linetype = 2) +
               geom_smooth(aes(match_day, 
                               avg_goal_per_day, 
                               color = league), se = FALSE) +
               labs(x = "Match Day",
                    y = "Average goal scored per match day") +
               theme_bw()
})

```
> Approaching the end of the season the number of goals scored in a match on avearge increase exponentially


### Average expected goal per game played across the years 

```{r}

# here is the same plat as above but the expected goal is used and we show them for all of the years. here we want to compared if there has been a change in trend across different years (Nicholas)

renderPlot({
  
  data_per_game1 |> filter(league %in% input$select_options) |> 
                    group_by(year, league, match_day) |>
                    summarise(avg_exp_goal = sum(exp_goal)/(n()/2)) |> 
                    ggplot() + geom_line(aes(match_day, 
                                             avg_exp_goal,
                                             color = league), 
                                         linetype = 6, 
                                         linewidth = 0.3) +
                    geom_smooth(aes(match_day, 
                                    avg_exp_goal, 
                                    color = league), 
                                se = FALSE) +
                    labs(x = "Match day",
                         y = "Average expected goal scored per match day") +
                    theme_bw() +
                    facet_wrap(~ year)
  })

              
```

> Along the years the expected goal per match-played increased torward the end of the season, supporting the idea that teams create more goal occasions.

###  Average passeges completed per goal scored during a match among the top 5 leagues

```{r}

# this Scatterplot compares the average of passages completed per goal with the average of goals per match played
# here again the trend line is shown to get an idea on the type of the trend. (Nicholas)

renderPlot({
   
  data_per_club |> filter(year == input$year_season, 
                          league %in% input$select_options) |> group_by(league) |>
                   ggplot() + 
                   geom_point(aes(scored/matches, deep/matches)) +
                   geom_smooth(aes(scored/matches, deep/matches)) + 
                   labs(x = "Average goal scored per match played by the team",
                        y = "Average passages completed per match played by the team ") +
                   theme_bw() +
                   facet_wrap(~league) 
  
  })


```
> There is an increasing trend between passages completed before scoring a goal and number of goals, meaming that scoring more goals requires play more with the ball.


column
-------------------------

### Expected goals scored and goals scored per teams winning the competition 

```{r}
# for the five teams that won the league in their respective country is shown a barplot that compares their average tot goal scored with the expected total goals scored. (Nicholas)

 renderPlot({
          
     data_per_club |> filter(position == 1, 
                             year == input$year_season, 
                             league %in% input$select_options) |> 
                      ggplot( aes(x = team)) + 
                      geom_bar(aes(y = exp_goal), position = "dodge", stat = "Identity") +
                      geom_point(aes(y = scored, color = "Goals scored"), shape = 17, size = 3) +
                      labs(x = "Winning teams for each league ",
                           y = "Expected total goal during the season ") +
                      scale_colour_manual("Total goals scored", breaks = "Goals scored", values = "red") +
                      theme_bw()
     
   })


```

> The winning teams scored more goals than the expected, indicating their ability to convert in goal shots


### Expected total points and effective points earned by the winning teams

```{r}

# the same as above but here we choose as metrics the total points earned and the total expected points (Nicholas)

renderPlot({
          
     data_per_club |> filter(position == 1, 
                             year == input$year_season, 
                             league %in% input$select_options) |> 
                      ggplot( aes(x = team)) + 
                      geom_bar(aes(y = exp_pts), position = "dodge", stat = "Identity") +
                      geom_point(aes(y = pts, color = "Points"), shape = 17, size = 3) + 
                      labs(x = "Winning teams for the top 5 leagues",
                           y = "Expected total points earned during a season ") +
                      scale_colour_manual("Total points", breaks = "Points", values = "blue") +
                      theme_bw()
                     
   })
  
```

> Teams that won the league had a higher ability to bring at home the points compared to their expected points, meaning that their efficiency in defense and attack made the difference

