---
title: "Football statistics dashboard"
output:
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
runtime: shiny
---

<style>
  .center-text {
    text-align: center;
  }
</style>

<div class="center-text">
  The best interactive dashboard on the market with statistics about football in the top 5 leagues
</div>


```{r setup, include = FALSE}
library(flexdashboard) 
library(tidyverse)
library(ggrepel)
library(DT)
library(shiny)
#knitr::opts_chunk$set(fig.width = 5)


#read the data
data1 <- read_csv(file = "data/understat_per_game.csv")
data2 <- read_csv(file = "data/understat.com.csv")

colnames_1 = c("league", "year", "stadium", "exp_goal", "exp_goal_against", 
               "no_pen_exp_goal", "no_pen_exp_goal_against", "deep", 
               "deep_allowed", "scored", "conceded", "exp_pts", "result", "date", 
               "wins", "draws", "loses", "pts", "no_pen_exp_goal_diff", "ppda_coef", 
               "ppda_att", "ppda_def", "oppda_coef", "oppda_att", "oppda_def", 
               "team", "exp_goal_diff", "exp_goal_against_diff", 
               "exp_pts_diff")

colnames(data1) = colnames_1

colnames_2 = c("league", "year", "position", "team", "matches", "wins", "draws", 
               "loses", "scored", "conceded", "pts", "exp_goal", "exp_goal_diff", 
               "no_pen_exp_goal", "exp_goal_against", "exp_goal_against_diff", 
               "no_pen_exp_goal_against", 
               "no_pen_exp_goal_diff", "ppda_coef", "oppda_coef", "deep", 
               "deep_allowed", "exp_pts", "exp_pts_diff")

colnames(data2) = colnames_2

data_per_game <- data1 |> 
  mutate("stadium" = as.factor(stadium), 
         "result" = as.factor(result), 
         "league" = as.factor(league),
         "pts" = as.factor(pts)) |> 
  filter(league != "RFPL")

data_per_club <- data2 |> 
  filter(league != "RFPL") |> 
  mutate("goal_diff" = scored - conceded)

## Check to see no duplicates and the pairing of opponent is done correctly on a 1-1 basis
length(data_per_game$league[duplicated(cbind(data_per_game$league, 
                                             data_per_game$year, 
                                             data_per_game$date, 
                                             data_per_game$scored, 
                                             data_per_game$conceded, 
                                             data_per_game$exp_goal_against, 
                                             data_per_game$exp_goal, 
                                             data_per_game$ppda_coef, 
                                             data_per_game$oppda_coef))])

## Link opponents to each respective match
temp_opponents = data_per_game |> select(league, year, date, scored, 
                                         conceded, exp_goal_against, exp_goal, 
                                         ppda_coef, oppda_coef, "opponent" = team)

data_per_game = data_per_game |> 
  left_join(temp_opponents, by = c("league" = "league", "year" = "year", 
                                   "date" = "date", "scored" = "conceded", 
                                   "conceded" = "scored", 
                                   "exp_goal_against" = "exp_goal", 
                                   "exp_goal" = "exp_goal_against", 
                                   "ppda_coef" = "oppda_coef", 
                                   "oppda_coef" = "ppda_coef"))


```





Team {data-icon="fa-futbol" data-orientation=rows}
==========================


Options {.sidebar}
------------------------

```{r}

selectInput("team_league", "Choose a competition:",
                c("Bundesliga" = "Bundesliga",
                  "Premier League" = "EPL",
                  "Serie A" = "Serie_A",
                  "La Liga" = "La_liga", 
                  "Ligue 1" = "Ligue_1" 
                  ))

selectInput("team_year", "Choose a season:",
                c("2014-2015" = "2014",
                  "2015-2016" = "2015",
                  "2016-2017" = "2016",
                  "2017-2018" = "2017", 
                  "2018-2019" = "2018", 
                  "2019-2020" = "2019"))

teams = reactive({sort(filter(data_per_club, league == input$team_league, year == input$team_year)$team)})

teams_opp = reactive({sort(filter(data_per_club, league == input$team_league, year == input$team_year, team != input$team_team)$team)})

renderUI({selectInput("team_team", "Choose a team:", teams())})

renderUI({selectInput("team_opp", 
                      "Choose a team for comparison:", 
                      teams_opp())})

```


tabset
-----------------------------------

<!-- ### tab1  -->

<!-- Here we can see the main statistics for the team (insert team name) and their historic results against team (insert opponents name) -->

<!-- row -->
<!-- --------------------------- -->

### Final Position
```{r fig.width=1, fig.height=0.5}

renderValueBox({
  position = filter(data_per_club, 
                    league == input$team_league, 
                    year == input$team_year, 
                    team == input$team_team)$position[1]
  
  box_icon = ""
  
  if(position == 1){
    box_icon = "ion-trophy"
  } else if(position <= 3){
    box_icon = "ion-ribbon-a"
  } else if(position <= 10){
    box_icon = "ion-android-happy"
  } else{
    box_icon = "ion-android-sad"
  }
  
  valueBox(value = position, icon = box_icon)
})

```

### Matches Won at Home
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "h")$wins)
  matches_won = sum(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "h")$wins)
  
  gauge(matches_won, min = 0, max = matches_played, gaugeSectors(
  success = c(2*matches_played/3, matches_played), warning = c(matches_played/3, 2*matches_played/3), danger = c(0, matches_played/3), colors = c("mediumseagreen", "palegoldenrod", "lightcoral")))
})

```
> Number of wins playing at home

### Matches Won Away
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "a")$wins)
  matches_won = sum(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "a")$wins)
  
  gauge(matches_won, min = 0, max = matches_played, gaugeSectors(
  success = c(2*matches_played/3, matches_played), warning = c(matches_played/3, 2*matches_played/3), danger = c(0, matches_played/3), colors = c("mediumseagreen", "palegoldenrod", "lightcoral")))
})

```
> Number of wins playing away

### Matches Lost at Home
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "h")$loses)
  matches_lost = sum(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "h")$loses)
  
  gauge(matches_lost, min = 0, max = matches_played, gaugeSectors(
  success = c(2*matches_played/3, matches_played), warning = c(matches_played/3, 2*matches_played/3), danger = c(0, matches_played/3), colors = c("lightcoral", "palegoldenrod", "mediumseagreen")))
})

```
> Number of loses playing at home

### Matches Lost Away
```{r fig.width=1, fig.height=1}

renderGauge({
  matches_played = length(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "a")$loses)
  matches_lost = sum(filter(data_per_game, league == input$team_league, year == input$team_year, team == input$team_team, stadium == "a")$loses)
  
  gauge(matches_lost, min = 0, max = matches_played, gaugeSectors(
  success = c(2*matches_played/3, matches_played), warning = c(matches_played/3, 2*matches_played/3), danger = c(0, matches_played/3), colors = c("lightcoral", "palegoldenrod", "mediumseagreen")))
})

```
> Number of loses playing away

row
--------------------------

### Season Results

```{r}

renderDataTable({data_per_game |> filter(league == input$team_league, 
                                     team == input$team_team, 
                                     year == input$team_year) |> 
    mutate(date = format(date, "%Y-%m-%d")) |> 
    mutate("home_team" = ifelse(stadium == "h", team, opponent)) |> 
    mutate("away_team" = ifelse(stadium == "a", team, opponent)) |> 
    mutate("home_goals" = ifelse(stadium == "h", scored, conceded)) |> 
    mutate("away_goals" = ifelse(stadium == "a", scored, conceded)) |> 
    select(date, home_team, home_goals, away_goals, away_team) |> 
    rename("Date" = "date", 
           "Home Team" = "home_team", 
           "Home Goals" = "home_goals", 
           "Away Goals" = "away_goals", 
           "Away Team" = "away_team")})

```
> Full team's results for the season

row
----------------------------------

### Head to Head 

```{r}

renderTable({data_per_game |> filter(league == input$team_league, 
                                     team == input$team_team, 
                                     opponent == input$team_opp) |> 
    mutate(date = format(date, "%Y-%m-%d")) |> 
    mutate("home_team" = ifelse(stadium == "h", team, opponent)) |> 
    mutate("away_team" = ifelse(stadium == "a", team, opponent)) |> 
    mutate("home_goals" = ifelse(stadium == "h", scored, conceded)) |> 
    mutate("away_goals" = ifelse(stadium == "a", scored, conceded)) |> 
    select(date, home_team, home_goals, away_goals, away_team) |> 
    rename("Date" = "date", 
           "Home Team" = "home_team", 
           "Home Goals" = "home_goals", 
           "Away Goals" = "away_goals", 
           "Away Team" = "away_team")}, digits = 0, align = "c")
  

```
> Historic results between the two teams in the competition

### Cummulative points comparison per match day

```{r}

renderPlot({data_per_game |> 
    filter((team == input$team_team | team == input$team_opp) & 
             year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    group_by(team, match) |> 
    summarize(pts = sum(pts)) |> 
    mutate(pts_cum = cumsum(pts)) |> 
    ggplot(mapping = aes(match, pts_cum, color = team)) + 
    geom_line() + 
    geom_point() + 
    labs(x = "Match Day", y = "Cummulative Points") + 
    theme_bw()})

```
> Comparison of performance of both teams throughout the season

row {.tabset .tabset-fade}
--------------------------

### Expected Goals and Actual Goals Scored throughout the Season

```{r}

renderPlot({data_per_game |> 
    filter(league == input$team_league, 
           team == input$team_team, 
           year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    ggplot() + 
    geom_line(mapping = aes(match, exp_goal, color = "Expected Goals"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, exp_goal, color = "Expected Goals"), 
              show.legend = TRUE) + 
    geom_line(mapping = aes(match, scored, color = "Goals Scored"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, scored, color = "Goals Scored"), 
              show.legend = TRUE) + 
    scale_color_manual("", breaks = c("Expected Goals", "Goals Scored"), 
                       values = c("red", "blue")) + 
    labs(x = "Match Day", y = "Goals / Expected Goals") + 
    theme_bw()})

```

### Expected Goals Against and Actual Goals Conceded throughout the Season

```{r}

renderPlot({data_per_game |> 
    filter(league == input$team_league, 
           team == input$team_team, 
           year == input$team_year) |> 
    group_by(team) |> 
    mutate("match" = c(1:n()), pts = as.numeric(as.character(pts))) |> 
    ungroup() |> 
    ggplot() + 
    geom_line(mapping = aes(match, exp_goal_against, 
                            color = "Expected Goals Against"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, exp_goal_against, 
                            color = "Expected Goals Against"), 
              show.legend = TRUE) + 
    geom_line(mapping = aes(match, conceded, color = "Goals Conceded"), 
              show.legend = TRUE) + 
    geom_point(mapping = aes(match, conceded, color = "Goals Conceded"), 
              show.legend = TRUE) + 
    scale_color_manual("", breaks = c("Expected Goals Against", 
                                                    "Goals Conceded"), 
                       values = c("red", "blue")) + 
    labs(x = "Match Day", y = "Goals / Expected Goals") + 
    theme_bw()})

```



National League {data-icon="fa-shield"}
==============================

Options {.sidebar}
------------------------

```{r}
selectInput("variable", "Choose a league:",
                c("Bundesliga" = "Bundesliga",
                  "Premier League" = "EPL",
                  "Serie A" = "Serie_A",
                  "La Liga" = "La_liga", 
                  "Ligue 1" = "Ligue_1"))
```


```{r}
selectInput("nation", "Choose a league:",
                c("Bundesliga" = "Bundesliga",
                  "Premier League" = "EPL",
                  "Serie A" = "Serie_A",
                  "La Liga" = "La_liga", 
                  "Ligue 1" = "Ligue_1"))


```

```{r}
selectInput("league", "Choose a league:",
                c("Bundesliga" = "Bundesliga",
                  "Premier League" = "EPL",
                  "Serie A" = "Serie_A",
                  "La Liga" = "La_liga", 
                  "Ligue 1" = "Ligue_1"))

```


column
----------------------------


### summary chart
```{r}



```





### table


```{r}


```


column  {.tabset}
-------------------------------

### chart 1

```{r}


```


### chart 2

```{r}

```

Comparison across leagues {data-orientation=rows data-icon="fa-globe"}
=================================


Options {.sidebar}
------------------------

```{r}
selectInput("year_season", "Choose the season:",
                c("2014/2015" = 2014,
                  "2015/2016" = 2015,
                  "2016/2017" = 2016,
                  "2017/2018" = 2017, 
                  "2018/2019" = 2018,
                  "2019/2020" = 2019))
```



```{r}
# Define the checkbox group input

checkboxGroupInput("select_options", "Select the leagues:", 
                   choices = c("Bundesliga", "EPL", "Serie_A","La_liga" ,"Ligue_1"),
                   selected = c("Bundesliga", "EPL", "Serie_A","La_liga" ,"Ligue_1"))


```


row  {data-orentation=columns data-height=250}
---------------------------------

### Winning teams for each league

```{r, fig.width=3}

renderTable({
      data_per_club |> filter(year == input$year_season, position == 1) |> 
      select(year,league, team:pts, exp_pts, exp_goal) }, digits = 0, align = "l", spacing = "s")

```


row 
----------------------------------

### Distribution of gol scored and total expected goal for each league

```{r}

 renderPlot({
      data_per_club |> filter(league %in% input$select_options) |> 
                                 ggplot() + 
                                 geom_boxplot(aes(league, exp_goal, color = "Exp Tot Goals" )) +
                                 geom_boxplot(aes(league, scored, color = "Tot goals scored")) +
                                 labs(x = "League",
                                      y = "Distribution of total goals scored") +
                                 scale_colour_manual("Exp goal per match", 
                                                     breaks = c("Exp Tot Goals", 
                                                                "Tot goals scored"), 
                                                     values = c("black","lightblue")) +
                                 facet_wrap(~ year) +
                                 theme_bw()
   
   
      })


```

row
-------------------------------

### Difference goal scored for the first 10th ranked teams within each league

```{r, fig.height=3}

data_per_game <-data_per_game |> mutate("date" = as.Date(date, "%Y-%m-%d"))

data_per_game1 <- data_per_game |> 
                  group_by(year, league, team) |> 
                  mutate("match_day" = row_number())



renderPlot({
  
  data_per_club |> filter(year == input$year_season, 
                          position <= 10, 
                          league %in% input$select_options) |> 
                     ggplot() + 
                     geom_line(aes(position, goal_diff, color = league)) + 
                     labs(x = "Teams ranked by final position",
                          y = "Difference between goals scored and conceded ",
                          color = "Top 5 leagues") +
                     geom_point(aes(position, goal_diff, color = league)) +
                     scale_x_continuous(breaks = seq(1,10, by = 2)) +
                     theme_bw()
      
  })                                  


```


### Average teams' pressure within each league


```{r}

renderPlot({
  data_per_club |> filter(year == input$year_season, 
                          league %in% input$select_options) |> 
                 group_by(league) |> summarise(avg_opp_pressure = mean(ppda_coef)) |>
                 ggplot() + 
                 geom_bar(aes(league, avg_opp_pressure), stat = "Identity") + 
                 labs(x = "Top five football leagues",
                      y = "Average teams' pressure per league ") +
                 theme_bw()
  
})
```


row {.tabset .tabset-fade}
---------------------------------

### Average expected goal and goal scored per game played in the Premier League

```{r}



renderPlot({
  
  data_per_game1 |> filter(year == input$year_season, 
                           league == "EPL") |> group_by(match_day) |> 
                  summarise(avg_goal_per_day = sum(scored)/10, 
                            avg_exp_goal = sum(exp_goal)/10) |>
                    ggplot() + 
                  geom_bar(aes(match_day, avg_goal_per_day), fill="grey", stat = "Identity") +    
                  geom_line(aes(match_day, avg_exp_goal, color = "Exp Goals per match"), 
                                       linewidth = 0.5, 
                                       linetype = "dashed", show.legend = TRUE) + 
                  labs(x = "Match day",
                       y = "Average goal scored per match day") + 
                  scale_colour_manual("Exp goal per match", breaks = "Exp Goals per match", values = "red") +
                  scale_x_continuous(breaks = seq(1, 38, by = 2)) +
                  theme_bw()
  
  
})

```
> Here we can see for each match-day the difference between expected goal and gol scored for the Premier League, 


### Average goal scored per match played in the top 5 leagues

```{r}

renderPlot({
  
  data_per_game1 |> filter(year == input$year_season, 
                           league %in% input$select_options) |> group_by(league, match_day) |> 
                    summarise(avg_goal_per_day = sum(scored)/10) |>
               ggplot() + 
               geom_line(aes(match_day, 
                             avg_goal_per_day, 
                             color = league), 
                             linetype = 2) +
               geom_smooth(aes(match_day, 
                               avg_goal_per_day, 
                               color = league), se = FALSE) +
               labs(x = "Match Day",
                    y = "Average goal scored per match day") +
               theme_bw()
})

```
> Approaching the end of the season the number of goals scored in a match on avearge increase exponentially


### Average expected goal per game played across the years 

```{r}

# try to change the intensity of of the lines so that it's easier to get the trend per league
# fix also the denominator in the division for the exact number of matches played

renderPlot({
  
  data_per_game1 |> filter(league %in% input$select_options) |> 
                    group_by(year, league, match_day) |>
                    summarise(avg_exp_goal = sum(exp_goal)/(n()/2)) |> 
                    ggplot() + geom_line(aes(match_day, 
                                             avg_exp_goal,
                                             color = league), 
                                         linetype = 6, 
                                         linewidth = 0.3) +
                    geom_smooth(aes(match_day, 
                                    avg_exp_goal, 
                                    color = league), 
                                se = FALSE) +
                    labs(x = "Match day",
                         y = "Average expected goal scored per match day") +
                    theme_bw() +
                    facet_wrap(~ year)
  })

              
```

> Along the years the expected goal per match-played increased torward the end of the season, supporting the idea that teams create more goal occasions.

###  Average passeges completed per goal scored during a match among the top 5 leagues

```{r}

renderPlot({
   
  data_per_club |> filter(year == input$year_season, 
                          league %in% input$select_options) |> group_by(league) |>
                   ggplot() + 
                   geom_point(aes(scored/matches, deep/matches)) +
                   geom_smooth(aes(scored/matches, deep/matches)) + 
                   labs(x = "Average goal scored per match played by the team",
                        y = "Average passages completed per match played by the team ") +
                   theme_bw() +
                   facet_wrap(~league) 
  
  })


```
> There is an increasing trend between passages completed before scoring a goal and number of goals, meaming that scoring more goals requires play more with the ball.


column
-------------------------

### Expected goals scored and goals scored per teams winning the competition 

```{r}

 renderPlot({
          
     data_per_club |> filter(position == 1, 
                             year == input$year_season, 
                             league %in% input$select_options) |> 
                      ggplot( aes(x = team)) + 
                      geom_bar(aes(y = exp_goal), position = "dodge", stat = "Identity") +
                      geom_point(aes(y = scored, color = "Goals scored"), shape = 17, size = 3) +
                      labs(x = "Winning teams for each league ",
                           y = "Expected total goal during the season ") +
                      scale_colour_manual("Total goals scored", breaks = "Goals scored", values = "red") +
                      theme_bw()
     
   })


```

> The winning teams scored more goals than the expected, indicating their ability to convert in goal shots


### Expected total points and effective points earned by the winning teams

```{r}

renderPlot({
          
     data_per_club |> filter(position == 1, 
                             year == input$year_season, 
                             league %in% input$select_options) |> 
                      ggplot( aes(x = team)) + 
                      geom_bar(aes(y = exp_pts), position = "dodge", stat = "Identity") +
                      geom_point(aes(y = pts, color = "Points"), shape = 17, size = 3) + 
                      labs(x = "Winning teams for the top 5 leagues",
                           y = "Expected total points earned during a season ") +
                      scale_colour_manual("Total points", breaks = "Points", values = "blue") +
                      theme_bw()
                     
   })
  
```

> Teams that won the league had a higher ability to bring at home the points compared to their expected points, meaning that their efficiency in defense and attack made the difference

